# Anchoring Survey Experiment

## Overview

This project examined the anchoring effect by testing whether different starting numbers influenced participants’ estimates of the height of the tallest redwood tree.

## What I Did

-   Collected survey data from two anchor conditions
-   Compared group estimates using summary statistics, visualizations, and a t-test
-   Cleaned the data by filtering extreme outliers and calculated an anchoring index

## Results

Participants who received the higher anchor gave much larger estimates on average. The results supported the idea that numerical anchors can strongly influence judgment.

## Tools Used

R, ggplot2, Hypothesis Testing, Data Cleaning

---
title: "Anchoring Survey Experiment"
author: "Joey Kent"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```

```{r}
library(rio)
library(mosaic)
library(tidyverse)
library(car)
library(dplyr)
library(summarytools)
library(psych)
library(writexl)
library(openxlsx)
library(ggplot2)
library(tidyr)
library(agricolae)

anchor <- read.csv("/Users/joeykent/Desktop/winter2026/eng301/anchoring_survey_project/anchoring_survey.csv")

View(anchor)
head(anchor)
str(anchor)
table(anchor$exp_guess)

```

```{r}
favstats(anchor$guess ~ anchor$group)
describeBy(anchor$guess, anchor$group)
```

```{r}
# Factor binary variables
anchor$group <- as.factor(anchor$group)

```

```{r}
ggplot(anchor, aes(x = group, y = guess, fill = group)) +
  geom_boxplot() +
  theme_minimal()

ggplot(anchor, aes(x = group, y = guess, fill = group)) +
  geom_point() +
  theme_minimal()

ggplot(anchor, aes(x = group, y = guess, fill = group)) +
  stat_summary(fun = mean, geom = 'bar', position = 'dodge') +
  theme_minimal()

qqPlot(anchor$guess ~ anchor$group)

```

```{r}
# Filter out guesses more than 20,000 feet
clean_anchor <- anchor %>%
  filter(guess < 10000)

View(clean_anchor)
favstats(clean_anchor$guess ~ clean_anchor$group)
describeBy(clean_anchor$guess, clean_anchor$group)
```

```{r}
# Graphs
ggplot(clean_anchor, aes(x = group, y = guess, fill = group)) +
  geom_boxplot() +
  theme_minimal()

ggplot(clean_anchor, aes(x = group, y = guess, fill = group)) +
  geom_point() +
  theme_minimal()

ggplot(clean_anchor, aes(x = group, y = guess, fill = group)) +
  stat_summary(fun = mean, geom = 'bar', position = 'dodge') +
  theme_minimal()

```

```{r}
cleaner_anchor <- anchor %>%
  filter(guess < 5000)

favstats(cleaner_anchor$guess ~ cleaner_anchor$group)
describeBy(cleaner_anchor$guess, cleaner_anchor$group)

View(cleaner_anchor)

ggplot(cleaner_anchor, aes(x = group, y = guess, fill = group)) +
  geom_boxplot() +
  theme_minimal()

ggplot(cleaner_anchor, aes(x = group, y = guess, fill = group)) +
  geom_point() +
  theme_minimal()

ggplot(cleaner_anchor, aes(x = group, y = guess, fill = group)) +
  stat_summary(fun = mean, geom = 'bar', position = 'dodge') +
  theme_minimal()
```

```{r}
t.test(cleaner_anchor$guess ~ cleaner_anchor$group, alternative = 'less')
t.test(cleaner_anchor$guess ~ cleaner_anchor$group, conf.level = .95)$conf.int
qqPlot(cleaner_anchor$guess ~ cleaner_anchor$group)
```

```{r}
# Calculating Anchoring Index

# Finding means
favstats(anchor$guess ~ anchor$group)
favstats(clean_anchor$guess ~ clean_anchor$group)
favstats(cleaner_anchor$guess ~ cleaner_anchor$group)

# Original data (anchor)
mean_low <- 274.57
mean_high <- 2493.45
low_anchor <- 180
high_anchor <- 1200

anchoring_index <- (mean_high - mean_low) / (high_anchor - low_anchor)

# Clean data (clean_anchor)
mean_low <- 274.57
mean_high <- 1479.19
low_anchor <- 180
high_anchor <- 1200

anchoring_index_clean <- (mean_high - mean_low) / (high_anchor - low_anchor)

# Cleaner data (cleaner_anchor)
mean_low <- 274.57
mean_high <- 1335.74
low_anchor <- 180
high_anchor <- 1200

anchoring_index_cleaner <- (mean_high - mean_low) / (high_anchor - low_anchor)

# Calling anchor indexes
anchoring_index
anchoring_index_clean # < 10000
anchoring_index_cleaner # < 5000
```