# Doubling-Back Aversion Study

## Overview

This project analyzed how people make decisions when switching tasks, focusing on the effect of losing prior progress.

## What I Did

-   Conducted a 2x2 experimental study
-   Analyzed switching behavior using logistic regression
-   Used ANOVA to examine perceived differences in task completion and route construal

## Results

Participants were less likely to switch tasks when prior progress was framed as being lost. The project helped show how framing can influence decision-making behavior.

## Tools Used

R, Statistical Modeling, Data Analysis

---
title: "Doubling-Back Aversion Study"
author: "Joey Kent"
format: html
editor: visual
---

```{r}
library(rio)
library(mosaic)
library(tidyverse)
library(car)
library(dplyr)
library(summarytools)
library(psych)
library(writexl)
library(openxlsx)
library(ggplot2)
library(tidyr)
library(agricolae)
library(effectsize)

cho <- read.csv("/Users/joeykent/Desktop/winter2026/experiencing_research/a_Winter_26_Cho_Doubling_Back_March_9_2026_20_14_copy.csv")
head(cho)
View(cho)
```

```{r}
old_cho <- cho %>%
  mutate(deletion = 1 - deletion,
         task_remain = 1 - task_remain)
View(old_cho)

# Making the decision to stay the value of 1
# cho$switch <- 1 - cho$switch

# Reverse coding construal
cho_reverse <- old_cho %>%
  mutate(strtovr = (7+1) - strtovr)

View(cho_reverse)

#Factoring
old_cho$deletion <- factor(old_cho$deletion)
old_cho$task_remain <- factor(old_cho$task_remain)

cho$deletion <- factor(cho$deletion)
cho$task_remain <- factor(cho$task_remain)

cho_reverse$deletion <- factor(cho_reverse$deletion)
cho_reverse$task_remain <- factor(cho_reverse$task_remain)
```

```{r}
# Considering Outliers for Route Time

# favstats(cho$nodel_75_doctortime)
# cho_clean <- cho %>%
#   filter(doctor_time < 2000)

# favstats(cho_clean$doctor_time)
```

```{r}
# Switch graph
ggplot(cho, aes(x = deletion, y = switch, fill = task_remain)) +
  stat_summary(
    fun = mean,
    geom = "bar",
    position = position_dodge(width = 0.9)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  scale_x_discrete(labels = c("1" = "Deletion", "0" = "No Deletion")) +
  scale_fill_discrete(labels = c("1" = "100% Task Remaining", "0" = "75% Task Remaining")) +
  scale_fill_manual(
  values = c("#729b79", "#bacdb0"),
  labels = c("75% Task Remaining", "100% Task Remaining")
  ) +
  labs(
    title = "Proportion Switching by Condition",
    x = "Deletion Condition",
    y = "Proportion Switched",
    fill = "Task Remaining"
  ) +
  theme_minimal()

# Perceived route length graph
ggplot(cho, aes(x = deletion, y = log_time_diff, fill = task_remain)) +
  stat_summary(
    fun = mean,
    geom = "bar",
    position = position_dodge(width = 0.9)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  scale_x_discrete(labels = c("1" = "Deletion", "0" = "No Deletion")) +
  scale_fill_discrete(labels = c("1" = "100% Task Remaining", "0" = "75% Task Remaining")) +
  scale_fill_manual(
  values = c("#729b79", "#bacdb0"),
  labels = c("75% Task Remaining", "100% Task Remaining")
  ) +
  labs(
    title = "Perceived Route Length by Condition",
    x = "Deletion Condition",
    y = "Mean Log Difference: Doctor Time - School Time",
    fill = "Task Remaining"
  ) +
  theme_minimal()

# Perceived route length construal
ggplot(old_cho, aes(x = deletion, y = strtovr, fill = task_remain)) +
  stat_summary(
    fun = mean,
    geom = "bar",
    position = position_dodge(width = 0.9)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  scale_x_discrete(labels = c("0" = "Deletion", "1" = "No Deletion")) +
  scale_fill_discrete(labels = c("0" = "100% Task Remaining", "1" = "75% Task Remaining")) +
  scale_fill_manual(
  values = c("#729b79", "#bacdb0"),
  labels = c("100% Task Remaining", "75% Task Remaining")
  ) +
  labs(
    title = "Perceived Route Construal by Condition",
    x = "Deletion Condition",
    y = "Mean Route Construal Rating",
    fill = "Task Remaining"
  ) +
  theme_minimal()

```

```{r}
# Running logistic regression
model <- glm(
  switch ~ deletion * task_remain,
  data = cho,
  family = binomial
)

summary(model)

# Getting odds ratio and confidence interval
# exp(coef(model)) # how to get odds ratio, but we use that in the code chunk below
exp(coef(model))
exp(cbind(OR = coef(model), confint(model)))
```

The z score for the deletion variable is 2.282 (significant), task_remaining variable is 0.829 (not significant), and the relationship between the two variables is 0.737 (not significant). For odds ratio and confidence interval, deletion variable is 1.972 \[1.10, 3.56\] (significant), task_remaining variable is 1.289 \[0.71, 2.36\] (not significant), and the relationships between the two variables is 1.370 \[0.59, 3.17\] (not significant)

```{r}
# Running 2x2 ANOVA test for Perceived Route Length
route_length_model <- aov(
  log_time_diff ~ deletion * task_remain, data = cho
  )
summary(route_length_model)

eta_squared(route_length_model)

# Normality Test
plot(route_length_model, which = 2)
shapiro.test(residuals(route_length_model))
```

```{r}
# Running 2x2 ANOVA test for Perceived Route Construal
route_construal_model <- aov(
  strtovr ~ deletion * task_remain, data = cho
  )
summary(route_construal_model)

eta_squared(route_construal_model)

# Normality Test
plot(route_construal_model, which = 2)
shapiro.test(residuals(route_construal_model))
```